---
title: "Weight Lifting Classification Problem"
output: html_document
---

## Preparation

Loading dependent libraries. 

Loading Training and Testing datasets for data processing.

```{r Preparation}
library(caret)
library(randomForest)
training <- read.csv("pml-training.csv",row.names=1,na.strings = "")
testing <- read.csv("pml-testing.csv",row.names=1,na.strings = "NA")
```

## Data Cleaning

Removing constant parameters that has no contribution to the output.

```{r Data Cleaning, echo=FALSE}
nsv <- nearZeroVar(training,saveMetrics=TRUE)
```

## Split Train/Test Datasets

Cleansing Training and Testing datasets with the input from Data Cleaning step.

```{r pressure, echo=FALSE}
training <- training[,!nsv$nzv]
testing <- testing[,!nsv$nzv]
```

# Removing contextual and incomplete variables

Removing variables that were not collected for the all the experiments.

Removing variables that did not have proper numerical values.

Removing contextual variables since the experiments do not have sensitivity to measurement timestamp.

```{r Removing contextual and incomplete variables, echo=FALSE}
training_filter_na <- training[,(colSums(is.na(training)) == 0)]
testing_filter_na <- testing[,(colSums(is.na(testing)) == 0)]

colRm_train <- c("user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","num_window")
colRm_test <- c("user_name","raw_timestamp_part_1","raw_timestamp_part_2","cvtd_timestamp","num_window","problem_id")

training_colRm <- training_filter_na[,!(names(training_filter_na) %in% colRm_train)]
testing_colRm <- testing_filter_na[,!(names(testing_filter_na) %in% colRm_test)]

dim(training_colRm)
dim(testing_colRm)

inTrain <- createDataPartition(y=training$classe, p=0.7, list=FALSE)
training_clean <- training_colRm[inTrain,]
validation_clean <- training_colRm[-inTrain,]
```

## Building Models

Random forest technique is used since this is a typical classificaion problem.

```{r Building Models, echo=FALSE}
set.seed(1234)
rfFit <- train(classe ~ ., method = "rf", data = training_clean, importance = T, trControl = trainControl(method = "cv", number = 4))
validation_pred <- predict(rfFit, newdata=validation_clean)
confusionMatrix(validation_pred,validation_clean$classe)

imp <- varImp(rfFit)$importance
varImpPlot(rfFit$finalModel, sort = TRUE, type = 1, pch = 19, col = 1, cex = 1, main = "Importance of the Predictors")
```

## Prediction of Test dataset

Applying random forest models on the test dataset.

```{r Prediction of Test dataset, echo=FALSE}
testing_pred <- predict(rfFit, newdata=testing_colRm)
```

```{r Predicted Result}
plot(testing_pred)
```